---
title: "Stock Price Forecasting Dashboard"
author: "Alex Alvarez"
date: "`r Sys.Date()`"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)

library(tidyverse)
library(quantmod)
library(forecast)
library(tseries)
library(DT)
library(plotly)
library(htmltools)
```

<style>
body {
  background: #f4f7fb;
  font-family: "Segoe UI", Arial, sans-serif;
}

.navbar {
  background: linear-gradient(90deg, #0f172a, #2563eb) !important;
}

.chart-title {
  background: white;
  border-left: 6px solid #2563eb;
  padding: 18px;
  border-radius: 14px;
  margin: 15px 0;
  box-shadow: 0 6px 18px rgba(0,0,0,0.08);
}

.metric-grid {
  display: flex;
  gap: 20px;
  margin: 20px 0;
}

.metric-card {
  flex: 1;
  background: linear-gradient(135deg,#0f172a,#2563eb);
  color: white;
  padding: 20px;
  border-radius: 16px;
  text-align: center;
}

.metric-card h2 {
  color: white;
  font-size: 32px;
  margin: 0;
}

.metric-card p {
  color: white;
  margin-top: 8px;
}
</style>

```{r data}
invisible(
  getSymbols(
    "NVDA",
    src = "yahoo",
    from = "2018-01-01"
  )
)

nvda <- data.frame(
  date = index(NVDA),
  price = as.numeric(Ad(NVDA))
)

nvda_ts <- ts(nvda$price)
stationarity_test <- adf.test(nvda_ts)
differences_needed <- ndiffs(nvda_ts)


arima_model <- auto.arima(nvda_ts)

forecast_nvda <- forecast(
  arima_model,
  h = 30
)

future_dates <- seq(
  from = max(nvda$date) + 1,
  by = "day",
  length.out = 60
)

future_dates <- future_dates[
  weekdays(future_dates) %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
][1:30]

forecast_df <- data.frame(
  Date = future_dates,
  Forecast = round(as.numeric(forecast_nvda$mean), 2),
  Lo80 = round(as.numeric(forecast_nvda$lower[, 1]), 2),
  Hi80 = round(as.numeric(forecast_nvda$upper[, 1]), 2)
)

latest_price <- round(tail(nvda$price, 1), 2)
tomorrow_price <- round(as.numeric(forecast_nvda$mean[1]), 2)
forecast_price <- round(as.numeric(forecast_nvda$mean[30]), 2)
expected_change <- round(((forecast_price - latest_price) / latest_price) * 100, 2)
```

Stock Price Chart
=====================================

Row {data-height=450}
-------------------------------------

### NVDA Stock Price Trend

<div class="chart-title">
<b>NVDA Stock Price Trend</b><br><br>
Adjusted closing price from 2018 to present. NVIDIA shows a major upward trend after 2023, likely reflecting AI-related market demand.
</div>

```{r price-chart}
plot_ly(
  data = nvda,
  x = ~date,
  y = ~price,
  type = "scatter",
  mode = "lines"
) %>%
  layout(
    xaxis = list(title = "Date"),
    yaxis = list(title = "Adjusted Close Price"),
    height = 350
  )
```

ARIMA Diagnostics
=====================================

Row {data-height=350}
-------------------------------------



### Stationarity Test



```{r stationarity-test}
data.frame(
  Metric = c("ADF p-value", "Differences Needed"),
  Value = c(
    round(stationarity_test$p.value, 4),
    differences_needed
  )
)







```

The Augmented Dickey-Fuller test returned a p-value of 0.9542, indicating the original NVDA price series is non-stationary. The model required 2 differences before fitting the ARIMA model.


### ARIMA Model


```{r}
data.frame(
  Model = "ARIMA(5,2,0)",
  AIC = round(AIC(arima_model),2),
  BIC = round(BIC(arima_model),2)
)

```

The selected model was ARIMA(5,2,0), meaning five autoregressive terms were used after applying second-order differencing. This specification was automatically selected based on information criteria (AIC).




Row {data-height=400}
-------------------------------------

### ACF Plot


```{r acf-plot}
nvda_diff <- diff(nvda_ts, differences = 2)


Acf(nvda_diff)
```



### PACF Plot
```{r pacf-plot}
Pacf(nvda_diff)
```


The ACF and PACF plots were generated using the second-order differenced series. After differencing, the autocorrelations are substantially reduced, indicating the data is closer to stationarity and more suitable for ARIMA modeling.




Residual Diagnostics
=====================================




```{r}




checkresiduals(arima_model)
```




The residual diagnostics evaluate whether the ARIMA model successfully captured the underlying patterns in NVDA stock prices. Ideally, residuals should behave like random noise with no significant autocorrelation remaining.

### Residual Diagnostics Summary

The Ljung-Box test was statistically significant (p < 0.05), indicating that some autocorrelation remains in the residuals. This suggests the ARIMA(5,2,0) model does not fully capture all patterns present in NVDA stock prices. Future model improvements may include alternative ARIMA specifications or additional explanatory variables.



Forecast Summary
=====================================

Row {data-height=220}
-------------------------------------

### Key Metrics

```{r metric-cards, results='asis'}
cat(sprintf(
'
<div class="metric-grid">

<div class="metric-card">
<h2>$%.2f</h2>
<p>Latest NVDA Price</p>
</div>


<div class="metric-card">
<h2>$%.2f</h2>
<p>Tomorrow Forecast</p>
</div>
<div class="metric-card">
<h2>$%.2f</h2>
<p>30-Day Forecast</p>
</div>

<div class="metric-card">
<h2>%.2f%%</h2>
<p>Expected Change</p>
</div>

</div>
',
latest_price,
tomorrow_price,
forecast_price,
expected_change
))
```



The 30-day forecast shows a sharp decline from the latest NVDA price. This result should be interpreted with caution because ARIMA is using historical price patterns only and does not account for earnings, news, AI demand, or broader market conditions. The residual diagnostics also suggest the model can be improved in a future version.


Row {data-height=430}
-------------------------------------

### ARIMA Forecast Chart

<div class="chart-title">
<b>ARIMA Forecast</b><br><br>
The ARIMA model forecasts NVDA's next 30 trading periods. The wider confidence intervals show that uncertainty increases over time.
</div>

```{r forecast-chart}
plot_ly(forecast_df, x = ~Date) %>%
  add_ribbons(
    ymin = ~Lo80,
    ymax = ~Hi80,
    name = "80% Confidence Interval",
    opacity = 0.2
  ) %>%
  add_lines(
    y = ~Forecast,
    name = "Forecast"
  ) %>%
  layout(
    xaxis = list(title = "Date"),
    yaxis = list(title = "Forecasted Price")
  )
```



### Forecast Interpretation

The ARIMA model forecasts a downward trend in NVDA prices over the next 30 trading periods. However, forecasts should be interpreted with caution because the model relies solely on historical price movements and does not incorporate company fundamentals, earnings information, or market news. The residual diagnostics indicate that additional model refinement may improve forecasting accuracy.

Forecast Table
=====================================

Row {data-height=600}
-------------------------------------

### 30-Day NVIDIA Stock Forecast

```{r forecast-table}
datatable(
  forecast_df,
  options = list(
    pageLength = 10,
    scrollX = TRUE
  ),
  caption = "30-Day NVIDIA Stock Forecast"
)
```
