---
title: "Stock Price Forecasting Dashboard"
author: "Alex Alvarez"
date: "`r Sys.Date()`"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
---







```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)
```

<style>
body {
  background: #f4f7fb;
  font-family: "Segoe UI", Arial, sans-serif;
}

.navbar {
  background: linear-gradient(90deg, #0f172a, #2563eb) !important;
}

.chart-title {
  background: white;
  border-left: 6px solid #2563eb;
  padding: 18px;
  border-radius: 14px;
  margin: 15px 0;
  box-shadow: 0 6px 18px rgba(0,0,0,0.08);
}


.metric-grid {
  display: flex;
  gap: 20px;
  margin: 20px 0;
}

.metric-card {
  flex: 1;
  background: linear-gradient(135deg,#0f172a,#2563eb);
  color: white;
  padding: 20px;
  border-radius: 16px;
  text-align: center;
}

.metric-card h2 {
  color: white;
  font-size: 32px;
  margin: 0;
}

.metric-card p {
  color: white;
  margin-top: 8px;
}
</style>

```{r}

library(tidyverse)
library(quantmod)
library(forecast)
library(tseries)
library(DT)
library(plotly)
library(htmltools)
```


```{r}
invisible(
  getSymbols(
    "NVDA",
    src = "yahoo",
    from = "2018-01-01"
  )
)


nvda <- data.frame(
  date = index(NVDA), 
  price = as.numeric(Ad(NVDA))
)


```




## Stock price chart



Row {data-height=450}
-------------------------------------

### NVDA Stock Price Trend

<div class="chart-title">
<b>NVDA Stock Price Trend</b><br><br>
Adjusted closing price from 2018 to present. NVIDIA shows a major upward trend after 2023, likely reflecting AI-related market demand.
</div>







```{r}
plot_ly(
  data = nvda,
  x = ~date,
  y = ~price,
  type = "scatter",
  mode = "lines"

)
```


Row {data-height=220}
-------------------------------------


## ARIMA Forecast
```{r}
nvda_ts <- ts(nvda$price)

arima_model <- auto.arima(nvda_ts)

forecast_nvda <- forecast(
  arima_model,
  h = 30
)

future_dates <- seq(
  from = max(nvda$date) + 1,
  by = "day",
  length.out = 60
)

future_dates <- future_dates[weekdays(future_dates) %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")][1:30]

forecast_df <- data.frame(
  Date = future_dates,
  Forecast = round(as.numeric(forecast_nvda$mean), 2),
  Lo80 = round(as.numeric(forecast_nvda$lower[,1]), 2),
  Hi80 = round(as.numeric(forecast_nvda$upper[,1]), 2)
)

latest_price <- round(tail(nvda$price, 1), 2)

forecast_price <- round(
  as.numeric(forecast_nvda$mean[30]),
  2
)

expected_change <- round(
  ((forecast_price - latest_price) / latest_price) * 100,
  2
)

```


### Latest Price






## Forecast Results



<div class="chart-title">

<b>Forecast Summary</b><br><br>

The ARIMA model forecasts NVDA's next 30 trading periods. The wider confidence intervals show that uncertainty increases over time.

</div>


```{r}


datatable(
  forecast_df,
  options = list(
    pageLength = 10,
    scrollX = TRUE
  ),
  caption = "30-Day NVIDIA Stock Forecast"
)
```




```{r}

latest_price <- round(tail(nvda$price, 1), 2)

forecast_price <- round(
  as.numeric(forecast_nvda$mean[30]),
  2
)

expected_change <- round(
  ((forecast_price - latest_price) / latest_price) * 100,
  2
)

```





```{r, echo=FALSE}
cat(sprintf(
'
<div class="metric-grid">

<div class="metric-card">
<h2>$%.2f</h2>
<p>Latest NVDA Price</p>
</div>

<div class="metric-card">
<h2>$%.2f</h2>
<p>30-Day Forecast</p>
</div>

<div class="metric-card">
<h2>%.2f%%</h2>
<p>Expected Change</p>
</div>

</div>
',
latest_price,
forecast_price,
expected_change
))

```

