---
title: 'Version 3: TSLA Advanced Machine Learning Forecasting'
author: "Alex Alvarez"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: flatly
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
  warning = FALSE,
  message = FALSE)
```




```{r}
library(tidyverse)
library(quantmod)
library(randomForest)
library(xgboost)
library(DT)
library(plotly)
library(zoo)
```

# Load TSLA Data


```{r}

invisible(
  getSymbols(
    "TSLA", 
    src = "yahoo",
    from = "2018-01-01"
                     )
          )
tsla <- data.frame(
  date = index(TSLA), 
  price = as.numeric(Ad(TSLA)), 
  volume = as.numeric(Vo(TSLA))
)



head(tsla)
```

# Data Summary

The dataset contains Tesla (TSLA) daily stock prices and trading volume obtained from Yahoo Finance. The adjusted closing price is used as the target variable for forecasting, while trading volume will be incorporated as a potential predictor in the machine learning models.

```{r}


summary(tsla)
```







### Dataset Characteristics





The summary statistics provide an overview of Tesla's stock price and trading volume over the sample period. These statistics help identify the range, central tendency, and variation in the data before feature engineering and model development.


```{r}

range(tsla$date)


nrow(tsla)
```


The dataset spans multiple years of daily trading activity and contains sufficient observations for machine learning model development and out-of-sample testing.




# Feature Engineering
Machine learning models require informative predictor variables. To improve forecasting performance, several features were engineered from Tesla's historical stock prices and trading volume.

These features capture recent price movements, trend behavior, volatility, and trading activity.
```{r}





tsla_ml <- tsla %>%
  mutate(
    lag1 = lag(price, 1),
    lag2 = lag(price, 2),
    lag5 = lag(price, 5),
    lag10 = lag(price, 10),

    ma5 = rollmean(price, 5,
                   fill = NA,
                   align = "right"),

    ma20 = rollmean(price, 20,
                    fill = NA,
                    align = "right"),

    daily_return = (price / lag(price, 1)) - 1,

    volatility_5 = rollapply(
      daily_return,
      width = 5,
      FUN = sd,
      fill = NA,
      align = "right"
    )
  ) %>%
  drop_na()
head(tsla_ml)
```








```{r}

tail(tsla_ml
     )
```



# Train Test Split


```{r}
set.seed(123)

train_size <- floor(0.8 * nrow(tsla_ml))

train <- tsla_ml[1:train_size, ]
test <- tsla_ml[(train_size + 1):nrow(tsla_ml), ]



cat("Total Observations:", nrow(tsla_ml), "\n")
cat("Training Observations:", nrow(train), "\n")
cat("Testing Observations:", nrow(test), "\n")
```




### Train-Test Split Results

The dataset was divided into two subsets:

- Training Set (80%) used to train the machine learning models.
- Testing Set (20%) used to evaluate model performance on unseen data.

This separation helps prevent overfitting and provides a realistic assessment of forecasting accuracy.


```{r}

nrow(train)
nrow(test)

```


# Linear Regression Model


Linear Regression serves as a baseline forecasting model. The model uses lagged prices, moving averages, trading volume, daily returns, and volatility measures to predict Tesla's stock price.



```{r}


lm_model <- lm(
  price ~ lag1 + lag2 + lag5 + lag10 +
    ma5 + ma20 +
    volume +
    daily_return +
    volatility_5,
  data = train
)

summary(lm_model)

```



The Linear Regression model achieved an R² of 0.9988, indicating excellent explanatory power. Daily returns, the 5-day moving average, and the previous day's stock price were among the strongest predictors based on their statistical significance and t-statistics. These results suggest that recent momentum and short-term trend indicators contain substantial information for forecasting Tesla stock prices. 


```{r}

lm_pred <- predict(
  lm_model,
  newdata = test
)

rmse_lm <- sqrt(
  mean((test$price - lm_pred)^2)
)

rmse_lm
```

# Random Forest Model


Random Forest is an ensemble machine learning algorithm that combines many decision trees to improve predictive performance. Unlike Linear Regression, Random Forest can capture nonlinear relationships and interactions among variables.


```{r}



rf_model <- randomForest(
  price ~ lag1 + lag2 + lag5 + lag10 +
    ma5 + ma20 +
    volume +
    daily_return +
    volatility_5,
  data = train,
  ntree = 500,
  mtry = 3
)

rf_model
```

```{r}


rf_pred <- predict(
  rf_model,
  newdata = test
)

rmse_rf <- sqrt(
  mean((test$price - rf_pred)^2)
)

rmse_rf
```


The Random Forest model achieved an RMSE of 36.14 on the testing dataset, substantially higher than the Linear Regression model's RMSE of 7.12. Despite explaining approximately 99.79% of the variance in the training data, the Random Forest model did not generalize well to unseen observations. This suggests that Tesla's stock price relationships may be captured more effectively by the linear and trend-based features used in the regression model.





# XGBoost Model

XGBoost is a gradient boosting algorithm that builds trees sequentially, with each new tree attempting to correct errors made by previous trees. It is widely used in predictive analytics and machine learning competitions.


```{r}



x_train <- as.matrix(
  train[, c(
    "lag1",
    "lag2",
    "lag5",
    "lag10",
    "ma5",
    "ma20",
    "volume",
    "daily_return",
    "volatility_5"
  )]
)





y_train <- train$price

x_test <- as.matrix(
  test[, c(
    "lag1",
    "lag2",
    "lag5",
    "lag10",
    "ma5",
    "ma20",
    "volume",
    "daily_return",
    "volatility_5"
  )]
)

y_test <- test$price

xgb_model <- xgboost(
  data = x_train,
  label = y_train,
  nrounds = 100,
  objective = "reg:squarederror",
  verbose = 0
)



```



```{r}






xgb_pred <- predict(
  xgb_model,
  x_test
)

rmse_xgb <- sqrt(
  mean((y_test - xgb_pred)^2)
)

rmse_xgb
```






# Model Comparison


```{r}



results <- data.frame(
  Model = c(
    "Linear Regression",
    "Random Forest",
    "XGBoost"
  ),
  RMSE = c(
    rmse_lm,
    rmse_rf,
    rmse_xgb
  )
)

results
```


Three forecasting models were evaluated using Tesla stock data: Linear Regression, Random Forest, and XGBoost.

Model performance was measured using Root Mean Squared Error (RMSE) on the testing dataset.

The Linear Regression model achieved the lowest RMSE (7.12), substantially outperforming Random Forest (36.14) and XGBoost (41.13). These findings suggest that Tesla stock prices can be effectively modeled using lagged prices, moving averages, trading volume, daily returns, and volatility measures within a linear framework.

Although machine learning methods are often expected to outperform traditional statistical models, the results demonstrate that increased model complexity does not necessarily lead to better forecasting performance. For this dataset, Linear Regression provided the most accurate predictions and remains the preferred forecasting model.



# TSLA Forecast


The Linear Regression model achieved the lowest RMSE and was selected as the final forecasting model. The most recent Tesla observations are used to generate an out-of-sample forecast.

```{r}
latest_obs <- tail(tsla_ml, 1)

next_day_forecast <- predict(
  lm_model,
  newdata = latest_obs
)

next_day_forecast
```



The final Linear Regression model was used to generate an out-of-sample forecast using the most recent Tesla observations.

The latest observed Tesla stock price was $406.43. The model predicted a next-trading-day price of approximately $400.51.

This forecast suggests a modest short-term decline of approximately $5.92, or 1.46%, relative to the latest observed price. Forecasts should be interpreted as statistical estimates rather than guarantees of future market performance.

