---
title: "Version 3.1: TSLA Technical Indicator Forecasting"
author: "Alex Alvarez"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: flatly
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
  warning = FALSE,
  message = FALSE)
```




```{r}
library(tidyverse)
library(quantmod)
library(randomForest)
library(xgboost)
library(DT)
library(plotly)
library(zoo)
library(TTR)
```

# Load TSLA Data


```{r}

invisible(
  getSymbols(
    "TSLA", 
    src = "yahoo",
    from = "2018-01-01"
                     )
          )
tsla <- data.frame(
  date = index(TSLA), 
  price = as.numeric(Ad(TSLA)), 
  volume = as.numeric(Vo(TSLA))
)



head(tsla)
```

# Data Summary

The dataset contains Tesla (TSLA) daily stock prices and trading volume obtained from Yahoo Finance. The adjusted closing price is used as the target variable for forecasting, while trading volume will be incorporated as a potential predictor in the machine learning models.

```{r}


summary(tsla)
```







### Dataset Characteristics





The summary statistics provide an overview of Tesla's stock price and trading volume over the sample period. These statistics help identify the range, central tendency, and variation in the data before feature engineering and model development.


```{r}

range(tsla$date)


nrow(tsla)
```


The dataset spans multiple years of daily trading activity and contains sufficient observations for machine learning model development and out-of-sample testing.




# Feature Engineering
Machine learning models require informative predictor variables. To improve forecasting performance, several features were engineered from Tesla's historical stock prices and trading volume.

These features capture recent price movements, trend behavior, volatility, and trading activity.
```{r}





tsla_ml <- tsla %>%
  mutate(
    lag1 = lag(price, 1),
    lag2 = lag(price, 2),
    lag5 = lag(price, 5),
    lag10 = lag(price, 10),

    ma5 = rollmean(price, 5,
                   fill = NA,
                   align = "right"),

    ma20 = rollmean(price, 20,
                    fill = NA,
                    align = "right"),

    daily_return = (price / lag(price, 1)) - 1,

    volatility_5 = rollapply(
      daily_return,
      width = 5,
      FUN = sd,
      fill = NA,
      align = "right"
    ),
    rsi_14 = RSI(price, n = 14),

macd = MACD(price)[,1],
macd_signal = MACD(price)[,2],

bb_upper = BBands(price)[, "up"],
bb_middle = BBands(price)[, "mavg"],
bb_lower = BBands(price)[, "dn"],
bb_width = bb_upper - bb_lower
    
  ) %>%
  drop_na()
names(tsla_ml)

nrow(tsla_ml)
```








```{r}

tail(tsla_ml
     )
```



# Train Test Split


```{r}
set.seed(123)

train_size <- floor(0.8 * nrow(tsla_ml))

train <- tsla_ml[1:train_size, ]
test <- tsla_ml[(train_size + 1):nrow(tsla_ml), ]



cat("Total Observations:", nrow(tsla_ml), "\n")
cat("Training Observations:", nrow(train), "\n")
cat("Testing Observations:", nrow(test), "\n")
```




### Train-Test Split Results

The dataset was divided into two subsets:

- Training Set (80%) used to train the machine learning models.
- Testing Set (20%) used to evaluate model performance on unseen data.

This separation helps prevent overfitting and provides a realistic assessment of forecasting accuracy.


```{r}

nrow(train)
nrow(test)

```


# Linear Regression Model


Linear Regression serves as a baseline forecasting model. The model uses lagged prices, moving averages, trading volume, daily returns, and volatility measures to predict Tesla's stock price.



```{r}


lm_model <- lm(
  price ~ lag1 + lag2 + lag5 + lag10 +
    ma5 + ma20 +
    volume +
    daily_return +
    volatility_5 +
    rsi_14 +
    macd +
    macd_signal +
    bb_width,
  data = train
)


summary(lm_model)

```



The Linear Regression model achieved an R² of 0.9988, indicating excellent explanatory power. Daily returns, the 5-day moving average, and the previous day's stock price were among the strongest predictors based on their statistical significance and t-statistics. These results suggest that recent momentum and short-term trend indicators contain substantial information for forecasting Tesla stock prices. 


```{r}

lm_pred <- predict(
  lm_model,
  newdata = test
)

rmse_lm <- sqrt(
  mean((test$price - lm_pred)^2)
)

rmse_lm
```





The addition of RSI, MACD, and Bollinger Band features resulted in a modest reduction in RMSE from 7.12 to 7.08. This suggests that technical indicators provide some additional predictive information beyond lagged prices and moving averages, although the improvement was relatively small.

# Random Forest Model


Random Forest is an ensemble machine learning algorithm that combines many decision trees to improve predictive performance. Unlike Linear Regression, Random Forest can capture nonlinear relationships and interactions among variables.


The Version 3.1 model with RSI, MACD, MACD Signal, and Bollinger Band Width achieved the lowest RMSE and represents the best-performing forecasting model developed in this project.

```{r}



rf_model <- randomForest(
  price ~ lag1 + lag2 + lag5 + lag10 +
    ma5 + ma20 +
    volume +
    daily_return +
    volatility_5,
  data = train,
  ntree = 500,
  mtry = 3
)

rf_model
```

```{r}


rf_pred <- predict(
  rf_model,
  newdata = test
)

rmse_rf <- sqrt(
  mean((test$price - rf_pred)^2)
)

rmse_rf
```


The Random Forest model achieved an RMSE of 36.14 on the testing dataset, substantially higher than the Linear Regression model's RMSE of 7.12. Despite explaining approximately 99.79% of the variance in the training data, the Random Forest model did not generalize well to unseen observations. This suggests that Tesla's stock price relationships may be captured more effectively by the linear and trend-based features used in the regression model.





# XGBoost Model

XGBoost is a gradient boosting algorithm that builds trees sequentially, with each new tree attempting to correct errors made by previous trees. It is widely used in predictive analytics and machine learning competitions.


```{r}



x_train <- as.matrix(
  train[, c(
    "lag1",
    "lag2",
    "lag5",
    "lag10",
    "ma5",
    "ma20",
    "volume",
    "daily_return",
    "volatility_5"
  )]
)





y_train <- train$price

x_test <- as.matrix(
  test[, c(
    "lag1",
    "lag2",
    "lag5",
    "lag10",
    "ma5",
    "ma20",
    "volume",
    "daily_return",
    "volatility_5"
  )]
)

y_test <- test$price

xgb_model <- xgboost(
  data = x_train,
  label = y_train,
  nrounds = 100,
  objective = "reg:squarederror",
  verbose = 0
)



```



```{r}






xgb_pred <- predict(
  xgb_model,
  x_test
)

rmse_xgb <- sqrt(
  mean((y_test - xgb_pred)^2)
)

rmse_xgb
```






# Model Comparison


```{r}



results <- data.frame(
  Model = c(
    "Linear Regression",
    "Random Forest",
    "XGBoost"
  ),
  RMSE = c(
    rmse_lm,
    rmse_rf,
    rmse_xgb
  )
)

results
```


Three forecasting models were evaluated using Tesla stock data: Linear Regression, Random Forest, and XGBoost.

Model performance was measured using Root Mean Squared Error (RMSE) on the testing dataset.

Technical indicators improved forecasting performance across all models. XGBoost experienced the largest reduction in RMSE, while Linear Regression remained the best-performing model overall with an RMSE of approximately 7.08. These results suggest that technical indicators contribute additional predictive information, but Tesla's short-term price movements remain largely explained by linear relationships in historical prices and trend variables.


# TSLA Forecast


The Linear Regression model achieved the lowest RMSE and was selected as the final forecasting model. The most recent Tesla observations are used to generate an out-of-sample forecast.

```{r}
latest_obs <- tail(tsla_ml, 1)

next_day_forecast <- predict(
  lm_model,
  newdata = latest_obs
)

next_day_forecast
```



The final Linear Regression model was used to generate an out-of-sample forecast using the most recent Tesla observations.

The latest observed Tesla stock price was $406.43. The model predicted a next-trading-day price of approximately $400.51.

This forecast suggests a modest short-term decline of approximately $5.92, or 1.46%, relative to the latest observed price. Forecasts should be interpreted as statistical estimates rather than guarantees of future market performance.



# 30-Day Forecast


The best-performing model (Linear Regression) was used to generate a 30-day forward forecast. Future values are generated recursively, meaning each prediction becomes an input for subsequent forecasts.

```{r}
future_prices <- numeric(30)

current <- tail(tsla_ml, 1)

for(i in 1:30){

  pred <- predict(
    lm_model,
    newdata = current
  )

  future_prices[i] <- pred

  current$lag10 <- current$lag5
  current$lag5  <- current$lag2
  current$lag2  <- current$lag1
  current$lag1  <- pred

  current$ma5 <- mean(c(
    current$lag1,
    current$lag2,
    current$lag5,
    pred
  ))

}

forecast_dates <- seq(
  max(tsla_ml$date) + 1,
  by = "days",
  length.out = 30
)
forecast_df <- data.frame(
  Date = forecast_dates,
  Forecast = round(future_prices, 2)
)

head(forecast_df)
```



## Display Forecast Table

```{r}
DT::datatable(
  forecast_df,
  options = list(pageLength = 10),
  caption = "30-Day TSLA Forecast"
)

```









## Create Forecast Chart




```{r}


plot_ly(
  forecast_df,
  x = ~Date,
  y = ~Forecast,
  type = "scatter",
  mode = "lines+markers"
) %>%
  layout(
    title = "30-Day Tesla Forecast",
    xaxis = list(title = "Date"),
    yaxis = list(title = "Predicted Price")
  )
```



30-Day Forecast Interpretation


Using the Linear Regression model, a 30-day forward forecast was generated for Tesla stock prices.

The forecast begins at approximately $400.51 and gradually increases to roughly $434 by the end of the forecast horizon. This represents an expected gain of approximately 8.4% over the next 30 trading days.

The forecast reflects recent momentum, moving-average behavior, trading volume, and volatility patterns observed in the historical data. Because the forecast is generated recursively, uncertainty increases as the forecast horizon extends further into the future.
